---
title: "Reproducible Research: Peer Assessment 1"
author: "Luis David Bedon Gomez"
date: "4 Mai 2017"
output: html_document
keep_md: true
---

## Introduction
The following report discuss the questions asked for the Course Project 1 of the reproducible Research Course of the Data Science Specialization from the Johns Hopkings University an Coursera, concearning the data about personal movement using activity monitoring devices.

## 1. Loading and preprocessing the data

### 1.1 Downloading and loading the data
The given data are stored in a zipped csv-File "activity.csv" and are loaded through the following routine:


```{r loading data}

setwd("~/Coursera/ReproducibleResearch")
library(dplyr)
library(ggplot2)

if(!file.exists("RepData_PeerAssessment1/activity.csv")){
  urldata<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  download.file(urldata,destfile = "activity.zip",method = "curl")
  unzip("activity.zip")
}

# NA-values are stored as "NA"
activity<-read.csv("activity.csv", stringsAsFactors = FALSE, na.strings = "NA")

```

### 1.2 Preprocessing the data

The loaded data were screened to understand the file using the following instruction:

```{r preprocessing}
str(activity)
```

Here we see the dimension of the file, the names of the columns, its classes and the content of the first rows.

The columns "steps" and "intervals"" are class "int", as we expect. Contrary, the column "date" is class "character" and we need it to be class "date". I will also
change the names of the columns by writing them with a capital letter.

The needed transformation goes as follows:

```{r as.Date}
activity$date<-as.Date(activity$date)
names(activity)<-c("Steps","Date","Interval")
class(activity$Date)
names(activity)
```
And now the class of "activity$step" is "date" as needed.





